Verifying Cache Coherence Protocols for an Infinite Family of Cache Hierarchies

### 2014-02-06 paper/Intuition.tex
stuff

We first reduce the proof of store atomicity to 4 invariants, presented in
Section~\ref{firstLevel}. These invariants in turn can be reduced to 4 more
invariants presented in Section~\ref{secondLevel}. The reason for choosing the 8
 invariants is because they are not specific to the MSI protocol;
any other protocol that obeys these invariants obeys Theorem
\fullref{storeAtomicity}. We finally present the key insight that we
discovered for proving the second-level invariants in the form of 3 invariants.
We believe that other invalidation-based protocols with inclusive caches (\ie{}
those cache hierarchies where a parent cache contains all the locations that
any of its children contains) like MESI, MOSI and MOESI also
satisfy these 3 fundamental properties, making their verification easier by reusing
portions of our MSI proof.
We informally show why this is the case using the examples of MESI,
MOSI and MOESI in Section~\ref{sec:Refinements}.
\adamc{...and maybe we don't, for space reasons. \texttt{:-)}}

@@

While we have proofs in Coq for each invariant, our discussion will be mainly
informal for brevity. A full description can be found at~\xxx{MURALI PUT
REFERENCE HERE}. \adamc{or just remind of the source code URL given earlier}

@@

\begin{figure}
\centering
\includegraphics[scale=.5]{resps}
\caption{An example showing a sequence of responses ping-ponging between a child and its parent
\adamc{For a camera-ready version, we should replace these bitmap images with vector images, for better printing.}}
\label{resps}
\end{figure}

>>>>

We first reduce the proof of store atomicity to 4 invariants, presented in
Section~\ref{firstLevel}. These invariants in turn can be reduced to 4 more
invariants presented in Section~\ref{secondLevel}. The reason for choosing the 8
 invariants is because they are not specific to the MSI protocol;
any other protocol that obeys these invariants obeys Theorem
\fullref{storeAtomicity}. We finally present the key insight that we
discovered for proving the second-level invariants in the form of 3 invariants.
We believe that other invalidation-based protocols with inclusive caches (\ie{}
those cache hierarchies where a parent cache contains all the locations that
any of its children contains) like MESI, MOSI and MOESI also
satisfy these 3 fundamental properties, making their verification easier by reusing
portions of our MSI proof.
We informally show why this is the case using the examples of MESI,
MOSI and MOESI in Section~\ref{sec:Refinements}.
%\adamc{...and maybe we don't, for space reasons. \texttt{:-)}}

@@

While we have proofs in Coq for each invariant, our discussion will be mainly
informal for brevity. As mentioned earlier, all the proofs can be found at~\xxx{MURALI PUT
REFERENCE HERE}. %\adamc{or just remind of the source code URL given earlier}

@@

\begin{figure}
\centering
\includegraphics[scale=.5]{resps}
\caption{An example showing a sequence of responses ping-ponging between a child and its parent}
%\adamc{For a camera-ready version, we should replace these bitmap images with vector images, for better printing.}}
\label{resps}
\end{figure}

### 2014-02-06 paper/System.tex
stuff

We have chosen as our concrete example a hierarchical MSI
protocol~\cite{MSI}. Logically, the system is organized in a tree
structure, with the root connected to the main memory and each leaf interfaced
with a processor (Figure~\ref{hier}). The system is parameterized by the shape
of the tree, and each node can have any number of children.
In the following, $\leaf(c)$ indicates
that a cache $c$ is a leaf, and $\parent(c,p)$ indicates that $p$ is the parent of
$c$.

@@

Input and output happen at the leaves in the form of memory requests
issued from and responses issued to the associated processors. These will be
represented by two mappings, one from $\mathbb{N}$ to requests and the other
from $\mathbb{N}$ to responses. For each leaf $c$, $c.\pos$ represents the
sequence number of the next request to process; it is set to $0$ initially.
\adamc{Add processors and main memory, labeled explicitly, to the diagram?}

@@

\begin{figure}
\begin{subfigure}{6.8cm}
\begin{tabular}{|lp{5.8cm}|}
\hline
\multicolumn{2}{|c|}{A message in \cpReq{} channel}\\
\hline
\from: & Current state of child\\
\myto: & Desired (higher) state of child\\
\addr: & Address associated with the request\\
\hline
\hline
\multicolumn{2}{|c|}{A message in \cpResp{} channel}\\
\hline
\from: & Current state of child\\
\myto: & Downgraded state of child\\
\addr: & Address associated with the request\\
\data: & Data associated with the address, if necessary\\
\hline
\end{tabular}
\end{subfigure}
\begin{subfigure}{5.4cm}
\begin{tabular}{|lp{4.4cm}|}
\hline
\multicolumn{2}{|c|}{A message in \pc{} channel}\\
\hline
\typ: & \Req{} (for requests) and \Resp{} (for responses)\\
\myto: & Desired (lower) directory state of a child for \Req{} and upgraded
directory state of a child for \Resp{}\\
\addr: & Address associated with the request\\
\data: & Data associated with the address if necessary (for \Resp{} only)\\
\hline
\end{tabular}
\end{subfigure}
\caption{Data types for messages between caches
\adamc{Names here don't match those in the preceding text.}}
\label{format}
\end{figure}

>>>>

We have chosen as our concrete example a hierarchical MSI
protocol. Logically, the system is organized in a tree
structure, with the root connected to the main memory and each leaf interfaced
with a processor (Figure~\ref{hier}). The system is parameterized by the shape
of the tree, and each node can have any number of children.
In the following, $\leaf(c)$ indicates
that a cache $c$ is a leaf, and $\parent(c,p)$ indicates that $p$ is the parent of
$c$.

@@

Input and output happen at the leaves in the form of memory requests
issued from and responses issued to the associated processors. These will be
represented by two mappings, one from $\mathbb{N}$ to requests and the other
from $\mathbb{N}$ to responses. For each leaf $c$, $c.\pos$ represents the
sequence number of the next request to process; it is set to $0$ initially.
%\adamc{Add processors and main memory, labeled explicitly, to the diagram?}

@@

\begin{figure}
\begin{subfigure}{6.8cm}
\begin{tabular}{|lp{5.8cm}|}
\hline
\multicolumn{2}{|c|}{A message in \cpReq{} channel}\\
\hline
\from: & Current state of child\\
\myto: & Desired (higher) state of child\\
\addr: & Address associated with the request\\
\hline
\hline
\multicolumn{2}{|c|}{A message in \cpResp{} channel}\\
\hline
\from: & Current state of child\\
\myto: & Downgraded state of child\\
\addr: & Address associated with the request\\
\data: & Data associated with the address, if necessary\\
\hline
\end{tabular}
\end{subfigure}
\begin{subfigure}{5.4cm}
\begin{tabular}{|lp{4.4cm}|}
\hline
\multicolumn{2}{|c|}{A message in \pc{} channel}\\
\hline
\typ: & \Req{} (for requests) and \Resp{} (for responses)\\
\myto: & Desired (lower) directory state of a child for \Req{} and upgraded
directory state of a child for \Resp{}\\
\addr: & Address associated with the request\\
\data: & Data associated with the address if necessary (for \Resp{} only)\\
\hline
\end{tabular}
\end{subfigure}
\caption{Data types for messages between caches}
%\adamc{Names here don't match those in the preceding text.}}
\label{format}
\end{figure}
